---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading and preprocessing the data
```{r data loading preprocess, echo = TRUE }
AMDdata<-read.csv(file="activity.csv")
```
## Calculating the total steps for each day
```{r histogram, echo=TRUE}
sumstepsperday<-tapply(AMDdata$steps,AMDdata$date,sum)
hist(sumstepsperday, main="Total steps per day", col="yellow",breaks = 10)
```

## What is mean total number of steps taken per day?
```{r mean/median, echo=TRUE}
meanstepsperday<-mean(sumstepsperday, na.rm=TRUE)
medianstepsperday<-median(sumstepsperday, na.rm=TRUE)
```
The mean is `r format(meanstepsperday,scientific=FALSE)`, and the median is `r medianstepsperday`

## What is the average daily activity pattern?
```{r activity pattern, echo=TRUE}
meanstepsinterval<-tapply(AMDdata$steps,AMDdata$interval,mean, na.rm=TRUE, simplify = TRUE)
meanstepsinterval<-as.data.frame(as.matrix(meanstepsinterval))
plot(row.names(meanstepsinterval),meanstepsinterval$V1, type="l", xlab="interval, 5min", ylab="Average steps", col="blue")
intervalwithmaxsteps<-row.names(meanstepsinterval)[which.max(meanstepsinterval$V1)]
```
Interval `r intervalwithmaxsteps`, on average across all the days in the dataset, contains the maximum number of steps

## Imputing missing values
```{r missing values, echo=TRUE}
narow<-nrow(AMDdata[is.na(AMDdata$steps),])
```
There are `r narow` missing values

###Strategy: replace the NA with averge steps for each interval
```{r Inputing missing values, echo=TRUE}
cleanAMDdata<-AMDdata
meanstepsinterval[,"interval"]<-rownames(meanstepsinterval)
cleanAMDdata<-merge(cleanAMDdata,meanstepsinterval,by="interval")
cleanAMDdata$steps[is.na(cleanAMDdata$steps)]<-cleanAMDdata$V1[is.na(cleanAMDdata$steps)]
cleanAMDdata$V1=NULL
```

##Calculate the total number steps, mean and median for each day 
```{r sum, echo=TRUE}
sumstepsperdaynew<-tapply(cleanAMDdata$steps,cleanAMDdata$date,sum)
hist(sumstepsperdaynew, main="Total steps per day", col="red",breaks = 10)
```

## What is mean total number of steps taken per day?
```{r new mean/median, echo=TRUE}
meanstepsperdaynew<-mean(sumstepsperdaynew )
medianstepsperdaynew<-median(sumstepsperdaynew)
```
The mean is `r format(meanstepsperdaynew,scientific=FALSE)`, and the median is `r format(medianstepsperdaynew,scientific=FALSE)`


## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays or weekends, echo=TRUE}
cleanAMDdata$Day <- weekdays(as.Date(as.character(cleanAMDdata$date)))
cleanAMDdata$Day[cleanAMDdata$Day=="Sunday"|cleanAMDdata$Day=="Saturday"]="weekend"
cleanAMDdata$Day[!(cleanAMDdata$Day=="weekend")]="weekday"
cleanAMDdata$Day <- as.factor(c("Weekend","Weekday"))
meanstepsnew<-with(cleanAMDdata, tapply(steps,list(interval, Day), mean, simplify = TRUE))
meanstepsnew<-as.data.frame(as.matrix(meanstepsnew))
```

```{r plotting, echo=TRUE}
plot(row.names(meanstepsnew), meanstepsnew$Weekday, type = "l", col = "green", main = "Weekdays", ylab="Average steps", xlab="Interval")
plot(row.names(meanstepsnew), meanstepsnew$Weekend, type = "l", col = "orange", main = "Weekends", ylab="Average steps", xlab="Interval")
```
